From Social Attention to Social Modelling in Robotics:\\ The Robotic Social Awareness Model

### 2016-07-01 paper.tex
The social hypothesis
WIP

\begin{abstract}
Abstract goes here
\end{abstract}

@@

A principled yet workable approach to social cognition for artificial intelligence

@@

@@

The hypothesis that we hereafter develop and turn into a cognitive model is the
following: mental representations are snapshots of awareness, awareness being
itself a label for the memory-mediated process of attention.

@@

@@

\begin{figure}
    \centering
        \resizebox{0.7\linewidth}{!}{%
        \begin{tikzpicture}[
                >=latex,
                every edge/.style={draw, very thick},
                every node/.style={font=\bf,align=center},
                input/.style={draw,circle, inner sep=0pt,minimum size=0.6cm}
            ]

@@

Grazino~\cite{graziano2013consciousness}: ``the mental machinery to model
someone else attentional state is the same as the one used for oneself.''
``In both case, the purpose is understanding and predicting one's behaviour.''

@@

@@

@@

@@

\subsection{Social Assessment}
\subsubsection{Gaze Direction}
\subsubsection{Facial and Body Expressions}
\subsubsection{Gestures}

>>>>

\begin{abstract}

My current perspective could be summarized as:

1- mental representations are snapshots of what we are aware of;

2- awareness is the label we conveniently put on the process of attention;

3- attention at time t is the label we put on the set of the activated units in
a (biased) associative memory network;

4- modelling others’ mental representations is taking snapshots of their own
current state of awareness;

5- modelling other’s state of awareness, ie, their current attentional process,
is mediated by one’s own attentional system, typically through joint attention
mechanisms;

6- Points 1 to 5 essentially refer to a 'phenomenal' awareness (a 'raw' inner
experience). 'Phenomenal' awareness can be turned into 'access' consciousness
(the abstract, cognitive ability to reflect on the inner experience);

7- In robots, 'access consciousness' can be mapped to symbolic, logic
representations.

(2) is Graziano's Attention Schemata Theory; (3) is based on Biased Competition
Model of Attention, Desimone and Duncan 1995; (6) comes from Block 1996.

As you hopefully see, this drafts a sub-symbolic/symbolic hybrid model of
cognition that puts (social) attention at the centre of social cognition.

Some of the main 'gaps' are:

- mental representations are likely not simply instantaneous 'snapshots' --
even though I could argue that they indirectly embed a notion of time and
history as they are snapshots of activated units in an associative memory
network;

- equating attention to a set of activated units in a memory network raises
questions regarding the nature of these units: physical entities? active
interaction modalities? ...? In particular, the right level of abstraction of
these units is not immediately clear: the spectrum is rather large, from raw
perceptual inputs 'à la Tony Morse' to high-level units like objects, joint
gazing, etc.

- the Biased Competition Model of Attention supports interesting bottom-up and
top-down biasing mechanisms: bottom-up is quite clear (if a unit is activated
longer/stronger, it biases the resulting attention to this unit. Top-down is
less clear (but also potentially very interesting as it closes the loop between
the sub-symbolic and symbolic models): how more abstract cognitive process can
influence the memory network to bias the attention process?

- the access to other's mental representations is mediated by one's own
attentional system: the consequences of this mediation are not immediately
clear

- I conveniently map the split between 'phenomenal consciousness' and 'access
consciousness' respectively to sub-symbolic and symbolic computational
structures. This need to be better evidenced, if only because the frontier
between phenomenal and access consciousness is all be clear (as pointed by
Graziano).

- It is not yet clear how the sub-symbolic and symbolic models of the different
agent co-exist: is each agent a different unit in the same associative network?
one associative memory per agent? what about the symbolic level?

Now, as we discussed many times, the model is of little interest if it does not
have some predictive power.

Here a few social behaviours that I would like a robot to exhibit if endowed
with such a model -- Not proper predictions, though.

- behavioural alignment -> surface alignment and/or global alignment -> the
original 'maze' task by Pickering and Garrod is certainly relevant here: every
time you play the same game with the same person, the interaction is more fluid
- ability to pass false-belief tasks, including non-physical, abstract ones, -
recursive awareness: being aware of being aware -- typically evidenced by being
able to describe/verbalize its own state of awareness - 'natural' turn-taking:
I 'just' know when it is my turn to act - 'natural' protodeclarative pointing:
I 'just' know when I really need to draw your attention on something

The word 'just' is clearly the key here: these capabilities should be natural
in the sense that they naturally follow from the model.

\end{abstract}

@@

A principled yet practical approach to social cognition for artificial intelligence

@@

\subsection{What is not adequately solved by current techniques?}

[give concrete examples of social situations that are not easily achieveable
with current (symbolic or not) approaches]

[this examples should be turned into predictions for what our proposal should be
able to achieve]

@@

The hypothesis that we hereafter develop and turn into a cognitive model is the
following: \textbf{mental representations are snapshots of awareness, awareness being
itself a label for the memory-mediated process of attention}.

This extends to social cognition: \textbf{modelling others' mental representations is
taking snapshots of their current state of awareness}. As we do not have direct
access to others' process of attention, it has to be mediated. We suggest that
\textbf{modelling other's state of awareness is mediated by one's own
attentional system, through joint attention mechanisms}.

This results in the following set of hypotheses:

\begin{itemize}
    \item \textbf{H1}: 
\end{itemize}

@@

analysis of behavioural alignment between partners (via
metrics like the recently proposed \emph{Individual Motor
Signature}~\cite{slowinski2016dynamic})

@@

\subsubsection{Memory-mediated Attention}

\emph{Attention} is modelled as the set of activated units in an associative
memory network.

Interplay between multi-modal social cues on one hand, and priming of previously activated physical
entities (objects, agents) and

Attention

\begin{figure}[h!]
    \centering
        \resizebox{0.7\linewidth}{!}{%
        \begin{tikzpicture}[
                >=latex,
                every edge/.style={draw, very thick},
                every node/.style={font=\bf,align=center},
                input/.style={draw,circle, inner sep=0pt,minimum size=0.6cm}
            ]

@@

Grazino~\cite{graziano2013consciousness}: ``the mental machinery to model
someone else attentional state is the same as the one used for oneself.''

``In both case, the purpose is understanding and predicting one's behaviour.''

@@

\subsection{Protodeclarative pointing}

Protoimperative vs protodeclarative pointing~\cite{baron1986perceptual}

@@

Picture ordering protocol, with \emph{mechanical}, \emph{behavioural} and
\emph{intentional} sequences~\cite{baron1986mechanical}

@@

\subsection{Adapted behaviour to the interacting agent}

A robot facing a baby, a 3 years old, a 5 years old, a 13 years old or an adult
should behave differently.

@@

\subsection{Social Assessment}

\subsubsection{Gaze Direction}
\subsubsection{Facial and Body Expressions}
\subsubsection{Gestures}
